---
title: "STAT332: Assignment 2 Question 5"
author: "Ashley Chui (21003047)"
output: pdf_document
---

\textbf{5.} Using the survey package with weight variable finalwt and properly accounting that SYC used a stratified sampling design,
give a point estimate and corresponding 95\% CI for each of the following:

```{r loading_data}
library("survey")
syc <- read.table("syc.txt", header = TRUE, sep = ",")

# Handle missing codes
syc$age[syc$age == 99] <- NA
syc$race[syc$race == 9] <- NA
syc$numarr[syc$numarr == 99] <- NA
syc$prviol[syc$prviol == 9] <- NA
syc$everdrug[syc$everdrug == 9] <- NA

# Survey design object
syc_des <- svydesign(
    ids = ~1, 
    strata = ~stratum,
    weights = ~finalwt,
    data = syc
)
df0 <- degf(syc_des)
```

(a) Estimate the average age of juveniles/young adults in custody.
```{r a, echo = TRUE}
a <- svymean(~age, syc_des, na.rm = TRUE)
confint(a, df = df0)
a
```
The point estimate for the average age of all juveniles/young adults in custody is 16.639 with a 95\% confidence interval $[16.5754, 16.7032]$.

(b) Estimate the mean number of prior arrests.
```{r b, echo = TRUE}
b <- svymean(~numarr, syc_des, na.rm = TRUE)
confint(b, df = df0)
b
```
The point estimate for the mean number of prior arrests is 8.930 with a 95\% confidence interval $[8.4181, 9.4414]$.

(c) Estimate the proportion of juveniles/young adults in custody that have used illegal drugs.
```{r c, echo = TRUE}
c <- svymean(~as.factor(everdrug), syc_des, na.rm =TRUE)
confint(c, df = df0)
c
```
The point estimate for the proportion of juveniles/young adults in custody that have used illegal drugs is 82.821\% with a 95\% confidence interval $[81.1052, 84.5370]$.\\

(d) Estimate the proportion of juveniles/young adults in custody that were previously arrested for a violent crime
```{r d, echo = TRUE}
d <- svymean(~prviol, syc_des, na.rm = TRUE)
confint(d, df = df0)
d
```
The estimated proportion of juveniles/young adults in custody that were previously arrested for a violent crime is 35.459\%. The 95% confidence interval for 
the proportion is $[33.3835\%, 37.5345\%]$.

(e) Estimate the proportion of juveniles/young adults in custody that have used illegal drugs and were previously arrested for a violent crime.
```{r e, echo = TRUE}
# create joint indicator variable
syc$vio_and_drug <- ifelse(syc$everdrug == 1 & syc$prviol == 1, 1, 0)
syc_des <- svydesign( # redefine design object
    ids = ~1, 
    strata = ~stratum,
    weights = ~finalwt,
    data = syc
)
e <- svymean(~vio_and_drug, syc_des, na.rm = TRUE)
confint(e, df = df0)
e
```
The estimated proportion of juveniles/young adults in custody that have used illegal drugs and were previously arrested for a violent crime is $31.389\%$ with
95\% confidence interval $[29.3927, 33.3851]$.\\

(f) Estimate the proportion of males amongst juveniles/young adults in custody.
```{r f, echo =TRUE}
f <- svymean(~factor(sex), syc_des, na.rm = TRUE)
confint(f, df = df0)
f
```
The estimated proportion of males amongst juveniles/young adults in custody is $93.119\%$ with 95\% confidence interval $[92.0510, 94.1876]$.\\

(g) Estimate the proportion of African Americans amongst juveniles/young adults in custody.
```{r g, echo = TRUE}
g <- svymean(~factor(race), syc_des, na.rm = TRUE)
confint(g, df = df0)
g
```
The estimated proportion of African Americans amongst juveniles/young adults in custody is $41.5303\%$ with 95\% confidence interval $[39.3550, 43.7055]$.\\

(h) Estimate the proportion of African American males amongst juveniles/young adults in custody.
```{r h, echo = TRUE}
syc$black_male <- ifelse(syc$race == 2 & syc$sex == 1, 1, 0)
syc_des <- svydesign( # redefine design object
    ids = ~1, 
    strata = ~stratum,
    weights = ~finalwt,
    data = syc
)
h <- svymean(~black_male, syc_des, na.rm = TRUE)
confint(h, df = df0)
h
```
The estimated proportion of African American males amongst juveniles/young adults in custody is $39.765\%$ with 95\% confidence interval $[37.6058, 41.9252]$.\

(i) From census data, we know that African-American males are about 7.6\% of the youth population
in the U.S. Using your answer to part (h), what can you conclude about racial bias when it comes to
youth in custody?

African-American males are overrepresented in youth custody by roughly 5 times their share of the general youth population.
This suggests substantial racial disparities and perhaps systemic bias in the youth justice system.