---
title: "STAT332: Assignment 1 Question 5"
author: "Ashley Chui (21003047)"
output: pdf_document
---

\textbf{5.} An instructor at the University of Florida conducted a 
small survey of the students in his introductory to statistics class. To this
end, he selected a simple random sample of 57 students from the 250 students 
in his class. The dataset file contains 11 variables, but this question is 
concerned with the 3 variables given in the table below.

```{r loading_data}
library("survey")

data <- read.table("classurv.txt", header = TRUE)

# remove missing GPA values coded as -9
data$GPA[data$GPA == -9] <- NA

# Define the survey design
dsgn <- svydesign(ids=~1, probs=c(57/350), fpc=rep(350, nrow(data)), data=data)
df0 <- degf(dsgn)
```

(a) Estimate the average GPA of all students
```{r a, echo = TRUE}
a <- svymean(~GPA, dsgn, na.rm=TRUE)
confint(a, df = df0)
a
```
The point estimate for the average GPA of all students is 1.11 with a 95\% confidence interval $[0.0347, 2.1860]$.

(b)-(c) Estimate average GPA of male/female students
```{r b, echo = TRUE}
b_c <- svyby(~GPA, ~gender, dsgn, svymean, na.rm = TRUE)
confint(b_c, df = df0)
b_c
```
The point estimate for the average GPA of male students is 3.049 with a 95\% confidence interval $[2.9227, 3.1756]$.\\
The point estimate for the average GPA of female students is 2.966 with a 95\% confidence interval $[2.8316, 3.1192]$.\\

(d) Estimate the proportions of male/female students
```{r d, echo = TRUE}
d <- svymean(~factor(gender), dsgn)
confint(d, df = df0)
d
```
The estimated proportion of male and female students is 40.35\% and 59.65\% respectively. The 95% confidence interval for 
the male proportion is $[28.33\%, 52.37\%], while for the female proportion it is $[47.63\%, 71.67\%]$.

(e) Estimate the proportions of students that are freshmen/sophomore/junior/senior/other
```{r e, echo = TRUE}
e_1 <- svyciprop(~I(year == 1), dsgn, method = "logit") # freshmen
e_1
e_2 <- svyciprop(~I(year == 2), dsgn, method = "logit") # sophomore
e_2
e_3 <- svyciprop(~I(year == 3), dsgn, method = "mean") # junior
e_3
e_4 <- svyciprop(~I(year == 4), dsgn, method = "logit") # senior
e_4
e_5 <- svyciprop(~I(year == 5), dsgn, method = "logit") # other
e_5
```

(f) Amongst freshmen, what are the proportions of male/female students
```{r f, echo =TRUE}
freshmen <- subset(dsgn, year == 1)
f_male <- svyciprop(~I(gender == 1), freshmen, method = "logit")
f_female <- svyciprop(~I(gender == 2), freshmen, method = "mean")
f_male
f_female
```

(g) Repeat (f), but for seniors
```{r g, echo = TRUE}
seniors <- subset(dsgn, year == 4)
g_male <- svyciprop(~I(gender == 1), seniors, method = "mean")
g_female <- svyciprop(~I(gender == 2), seniors, method = "mean")
g_male
g_female
```

(h) Amongst the top performers (i.e. GPA >= 3.2), what are the proportions of male/female students
```{r h, echo = TRUE}
top_perf <- subset(dsgn, GPA >= 3.2)
h_male <- svyciprop(~I(gender == 1), top_perf, method = "logit")
h_female <- svyciprop(~I(gender == 2), top_perf, method = "logit")
h_male
h_female
```